### 2022-08-14 osseu_2022/presentation.tex
OSSEU 2022

\documentclass[usenames,dvipsnames, 18pt, compress, aspectratio=169]{beamer}

% can be compiled by xelatex -shell-escape presentation.tex
% lualatex -shell-escape presentation.tex

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{xcolor}
\usepackage[document]{ragged2e}
\usepackage[export]{adjustbox}
\usepackage{fontawesome}
\usepackage{enumitem}
\usepackage{minted}
\usemintedstyle{tango}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{
    shapes,
    positioning
}

\usepackage[nott]{inconsolata}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{verbatim}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\usepackage{datetime}
\usepackage{hyperref}
\usepackage{forloop}
\usepackage{csquotes}

\usepackage{tcolorbox}
\usepackage{tabularx}
\usepackage{array}
\usepackage{colortbl}
\tcbuselibrary{skins}

\usetikzlibrary{shapes,arrows,positioning}
\graphicspath{{images}}

\def\twitter{{\FA \faTwitter}}
\def\github{{\FA \faGithub}}
\def\email{{\FA \faEnvelope}}

\renewcommand{\ttdefault}{pcr}
%\newfontfamily{\ttfamily}{SourceCodePro-Regular}

\usefonttheme{professionalfonts} % using non standard fonts for beamer
\usefonttheme{serif} % default family is serif
\usepackage{fontspec}
\setmainfont{Liberation Sans}
\newfontfamily\ExtraLight{Liberation Sans}
\newfontfamily\Light{Liberation Sans}
\newfontfamily\Book{Liberation Sans}
\newfontfamily\Medium{Liberation Sans}
\setmonofont{FiraCode-VF}

\makeatletter
\newcommand\HUGE{\@setfontsize\Huge{32}{41}}
\makeatother

\newcommand\AtPagemyUpperLeft[1]{\AtPageLowerLeft{%
\put(\LenToUnit{0.85\paperwidth},\LenToUnit{0.05\paperheight}){#1}}}

\newcommand\AtPagemyUpperTop[1]{\AtPageLowerLeft{%
\put(\LenToUnit{0.42\paperwidth},\LenToUnit{0.90\paperheight}){#1}}}

\renewcommand{\ULthickness}{2.0pt}

\definecolor{links}{HTML}{0099FF}
\hypersetup{colorlinks, linkcolor=, urlcolor=links}
\definecolor{title}{HTML}{ee0000}

\setbeamerfont{section title}{family=\Book, size=\Huge, shape=\normalfont}
\setbeamerfont{frametitle}{family=\Book, size=\large, shape=\normalfont}
\setbeamerfont{title}{family=\Book, size=\Large, shape=\normalfont}
\setbeamerfont{subtitle}{size=\small}
\setbeamerfont{author}{family=\ExtraLight, size=\footnotesize}

\setbeamercolor{frametitle}{fg=black}
\setbeamertemplate{frametitle}[default][center]
\setbeamerfont{frametitle}{size=\Large, series=\bfseries}

\setbeamersize{text margin left=0mm,text margin right=0mm}

\setbeamertemplate{navigation symbols}{}
\beamertemplatenavigationsymbolsempty
\pagenumbering{gobble}

\setbeamertemplate{title page}
{

  \vspace*{2.1cm}
  \hspace{7.0cm}
  \begin{minipage}[b][\paperheight]{0.5\textwidth}
  \begin{center}

    \ifx\inserttitle\@empty\else
    {{% \inserttitle is nonempty
      \raggedright%
      %\linespread{1.0}%
      \usebeamerfont{title}%
      \usebeamercolor[fg]{title}%
      %\vspace*{1.3em}
      \if@noSmallCapitals%
        \inserttitle%
      \else%
        \scshape{\color{title} \textbf{\begin{flushleft}\inserttitle\end{flushleft}}}%
      \fi%
      \vspace*{0.3em}
    }}
    \fi

    \vspace*{0.5em}%

    \ifx\insertsubtitle\@empty\else
    {{% \insertsubtitle is nonempty
      \usebeamerfont{subtitle}%
      \usebeamercolor[fg]{subtitle}%
      {\color{black} \insertsubtitle}%
      \vspace*{3.0em}%
    }}
    \fi

    \vspace*{1.0em}%

    \usebeamerfont{author}%
    \usebeamercolor[fg]{author}%
    {\begin{flushleft}\color{black} \insertauthor\end{flushleft}}%

    %\vspace*{1.5em}
    \fontsize{8pt}{10}\selectfont
    {\begin{flushleft}\color{black} 14-09-2022\end{flushleft}}%

    \vfill
    \vspace*{2em}
  \end{center}
  \end{minipage}
}

\setbeamertemplate{section page}
{
  \vspace{2em}
  \centering
  \begin{minipage}{22em}
    \usebeamercolor[fg]{section title}
    \usebeamerfont{section title}
    {\color{black} \insertsectionhead\\[-1ex]}
  \end{minipage}
  \par
}

\setbeamertemplate{footline}
{
\begin{beamercolorbox}[wd=\textwidth,ht=3ex,dp=3ex,leftskip=0.3cm,rightskip=0.3cm]{structure}
  \usebeamerfont{page number in head/foot}
  \insertframenumber
\end{beamercolorbox}
}

\title{Performance Insights into eBPF Step by Step}
\subtitle{}
\date{\today}
\author{Dmitrii Dolgov\\ Senior Software Engineer}
\institute{}

\tikzset{app/.style={
    draw,
    fill=blue!30,
    minimum height=2cm,
    minimum width=2cm,
    rounded corners=.55cm,
}}

\tikzset{kernel/.style={
    draw,
    fill=green!30,
    minimum height=2cm,
    minimum width=2cm,
    rounded corners=.55cm,
}}

\tikzset{bpf/.style={
    draw,
    fill=red!30,
    minimum height=2cm,
    minimum width=2cm,
    rounded corners=.55cm,
}}

\tikzset{connection/.style={
    draw=none
}}

\tikzset{arrow/.style={
    ->,>=stealth,
    line width=0.2mm,
    color=red!50,
}}

\begin{document}
{
  \usebackgroundtemplate{\includegraphics[width=\paperwidth]{title.png}}%
  \fontsize{17pt}{18}\selectfont
  \maketitle
}

\AddToShipoutPictureBG{
  \AtPagemyUpperLeft{{\includegraphics[width=2.0cm,keepaspectratio]{logo.png}}}
}

\setbeamertemplate{background canvas}{}

\fontsize{17pt}{18}\selectfont

\begin{frame}[fragile]{}
    \frametitle{}

    \begin{center}
        \begin{tikzpicture}[
                align=center,
                node distance=0.0cm
        ]
            \node[app] (app) {\textbf{App}};
            \node[connection, right=0.5cm of app] (plus1) {\textbf{+}};

            \node[kernel, right=0.5cm of plus1] (kernel) {\textbf{Kernel}};
            \node[connection, right=0.5cm of kernel, color=white, draw opacity=0, line width=0] (plus2) {\textbf{+}};

            \node[bpf, right=0.5cm of plus2, color=white, draw opacity=0] (bpf) {\textbf{BPF ?}};
            \node[connection, draw=none, color=white, draw opacity=0, line width=0, above=1cm of kernel] (hot-path) {Hot Path};
            \draw[arrow, draw opacity=0, line width=0, color=white, draw=none] (hot-path) to [out=60,in=90] (bpf);

        \end{tikzpicture}
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{}

    \begin{center}
        \begin{tikzpicture}[
                align=center,
                node distance=0.0cm
        ]
            \node[app] (app) {\textbf{App}};
            \node[connection, right=0.5cm of app] (plus1) {\textbf{+}};

            \node[kernel, right=0.5cm of plus1] (kernel) {\textbf{Kernel}};
            \node[connection, right=0.5cm of kernel] (plus2) {\textbf{+}};

            \node[bpf, right=0.5cm of plus2] (bpf) {\textbf{BPF ?}};
            \node[connection, above=1cm of kernel] (hot-path) {Hot Path};
            \draw[arrow] (hot-path) to [out=60,in=90] (bpf);

        \end{tikzpicture}
    \end{center}
\end{frame}

\fontsize{26pt}{26}\selectfont
\begin{frame}[fragile]{}
    \frametitle{}

    \begin{center}
        \textbf{How to measure?}
    \end{center}
\end{frame}

\fontsize{17pt}{18}\selectfont

\begin{frame}[fragile]{}
    \frametitle{Global kernel stats}

    \begin{center}
        \begin{minted}[fontsize=\Large]{shell-session}
    $ sysctl -w kernel.bpf_stats_enabled=1
    $ bpftool prog
        \end{minted}
        \vspace{0.5cm}
        \begin{minted}[fontsize=\Large]{shell-session}
    379: raw_tracepoint [...]
    run_time_ns 35875602162 run_cnt 160512637
        \end{minted}
    \end{center}
\end{frame}
\note{
    Pros: easy
    Cons: global
}

\begin{frame}[fragile]{}
    \frametitle{Good old printk}

    \begin{center}
        \begin{minted}[fontsize=\Large]{c}
    // somewhere inside your BPF prog
    bpf_trace_printk("Timestamp: %lld", ts);
        \end{minted}
        \vspace{0.5cm}
        \begin{minted}[fontsize=\Large]{shell-session}
    $ cat /sys/kernel/debug/tracing/trace_pipe
    $ bpftool prog tracelog
        \end{minted}
    \end{center}
\end{frame}
\note{
    Pros: easy, flexible
    Cons: slow, too much data
}

\begin{frame}[fragile]{}
    \frametitle{Probes before/after}

    \begin{center}
        \begin{minted}[fontsize=\Large]{shell}
    kfunc:do_syscall_64 /comm == "prog"/
    { @entry[kstack] = count(); }

    kretfunc:do_syscall_64 /comm == "prog"/
    { @return[kstack] = count(); }
        \end{minted}
    \end{center}
\end{frame}
\note{
    Pros: fast
    Cons: no ordering between progs makes it hard to find a proper point to
    attach to, maxactive, variance not relevant to the bpf prog.

    Not necessarily full syscall, maybe just before/after bpf prog.
}

\begin{frame}[fragile]{}
    \frametitle{Probes before/after}

    \begin{center}
        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large]{c}
    SEC("fentry/XXX")
    int BPF_PROG(fentry_XXX)
    {
        //...
    }
               \end{minted}
           \end{column}

           \begin{column}{0.5\textwidth}
               \begin{minted}[fontsize=\large]{c}
    SEC("fexit/XXX")
    int BPF_PROG(fentry_XXX)
    {
        //...
    }
                \end{minted}
            \end{column}
        \end{columns}
    \end{center}
\end{frame}
\note{
    Pros: fast, only target progs
    Cons: requires BTF
}

\begin{frame}[fragile]{}
    \frametitle{Probes before/after}

    \begin{center}
        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large, escapeinside=||]{asm}
  |\colorbox{red!20}{\textbf{nopl   0x0(%rax,%rax,1)}}|
   xchg   %ax,%ax
   push   %rbp
   mov    %rsp,%rbp
   sub    $0x20,%rsp
               \end{minted}
           \end{column}

           \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large, escapeinside=||]{asm}
  |\colorbox{red!20}{\textbf{callq  0xffffffffffe0096c}}|
   xchg   %ax,%ax
   push   %rbp
   mov    %rsp,%rbp
   sub    $0x20,%rsp
                \end{minted}
            \end{column}
        \end{columns}
    \end{center}
\end{frame}
\note{
    In this way bpftool profile & perf stat -b work
}

\begin{frame}[fragile]{}
    \frametitle{Profiling}

    \begin{center}
        \begin{minted}[fontsize=\large]{bash}
  Percent | uops_retired.stall_cycles
          :
          : if (duration_ns < min_duration_ns)
     0.00 :    9f:movabs $0xffffc9000009e000,%rdi
     0.00 :    a9:mov    0x0(%rdi),%rsi
          :
          : e = bpf_ringbuf_reserve(...)
    21.74 :    ad:movabs $0xffff888103e70e00,%rdi
     0.00 :    b7:mov    $0xa8,%esi
     0.00 :    bc:xor    %edx,%edx
     0.00 :    be:callq  0xffffffffc0f9fbb8
        \end{minted}
    \end{center}
\end{frame}
\note{
    Few samples, only an example. Uops_retired.stall_cycles = Cycles without
    actually retired uops.  BTF is needed, not only as debugging information,
    but also as perf facilitation.
}

\fontsize{26pt}{26}\selectfont
\begin{frame}[fragile]{}
    \frametitle{}

    \begin{center}
        \textbf{How to improve?}
    \end{center}
\end{frame}

\fontsize{17pt}{18}\selectfont

\begin{frame}[fragile]{}
    \frametitle{BPF Instruction Set}

    \begin{center}
        \begin{minted}[fontsize=\large]{c}
                       if (pid < ts)
        \end{minted}

        \vspace{0.5cm}

        \begin{columns}
            \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large, escapeinside=||]{asm}
    ; if (pid < ts)
      cmp    %rsi,%rdi
      jae    0x0000000000000068
               \end{minted}

           \end{column}

           \begin{column}{0.5\textwidth}
                \begin{minted}[fontsize=\large, escapeinside=||]{asm}
; if (pid < ts)
  cmp    %rdi,%rsi
  jbe    0x0000000000000065
                \end{minted}

            \end{column}
        \end{columns}

    \vspace{2.0cm}
    \href{https://pchaigno.github.io/bpf/2021/10/20/ebpf-instruction-sets.html}
         {\color{links}\fontsize{10pt}{0}\selectfont eBPF Instruction Sets}
    \end{center}
\end{frame}
\note{
    Influences size of the program and final instructions (more workarounds for
    v1 here vs v2).  Wors with -mcpu=v1/v2/v3 -mattr=+alu32 (for me ‘probe’
    wasn’t working) via llc. With clang use -mllvm -mcpu …
}

\begin{frame}[fragile]{}
    \frametitle{BPF 2 BPF}

    \begin{center}
        \begin{minted}[fontsize=\large]{c}
    #ifndef __inline
    # define __inline                         \
       inline __attribute__((always_inline))
    #endif

    static __inline int test_bpf2bpf(void) {}
        \end{minted}

        \vspace{0.5cm}

        \begin{minted}[fontsize=\large]{asm}
    # 0xffffffffc1513a68:
      nopl   0x0(%rax,%rax,1)
      xor    %eax,%eax
      push   %rbp
    # [...]
        \end{minted}
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{BPF 2 BPF}

    \begin{center}
        \begin{minted}[fontsize=\large]{c}
    static int test_bpf2bpf(void) { }
        \end{minted}

        \vspace{0.5cm}

        \begin{minted}[fontsize=\large]{asm}
    # 0xffffffffc15b810c:
      nopl   0x0(%rax,%rax,1)
      xor    %eax,%eax
    # [...]
      callq  0x0000000000002370

    # 0xffffffffc15ba47c:
      nopl   0x0(%rax,%rax,1)
      xchg   %ax,%ax
    # [...]
        \end{minted}
    \end{center}
\end{frame}
\note{
    Works similar to bpf helpers.
    The calling convention known from BPF helper function applies to BPF to BPF
    calls just as well, meaning r1 up to r5 are for passing arguments to the
    callee and the result is returned in r0. r1 to r5 are scratch registers
    whereas r6 to r9 preserved across calls the usual way. The maximum number
    of nesting calls respectively allowed call frames is 8. A caller can pass
    pointers (e.g. to the caller’s stack frame) down to the callee, but never
    vice versa.

    BPF JIT compilers emit separate images for each function body and later fix
    up the function call addresses in the image in a final JIT pass. This has
    proven to require minimal changes to the JITs in that they can treat BPF to
    BPF calls as conventional BPF helper calls.
}

\fontsize{26pt}{26}\selectfont
\begin{frame}[fragile]{}
    \frametitle{}

    \begin{center}
        \textbf{How to improve in the future?}
    \end{center}
\end{frame}

\fontsize{17pt}{18}\selectfont

\begin{frame}[fragile]{}
    \frametitle{Map batch operations}

    \begin{center}
        \begin{minted}[fontsize=\Large]{c}
    BPF_MAP_LOOKUP_BATCH
    BPF_MAP_LOOKUP_AND_DELETE_BATCH
    BPF_MAP_UPDATE_BATCH
    BPF_MAP_DELETE_BATCH
        \end{minted}
    \end{center}
\end{frame}
\note{
    aa2e93b8e58e18442edfb2427446732415bc215e
    cb4d03ab499d4c040f4ab6fd4389d2b49f42b5a5
    + Libbpf support
}

\begin{frame}[fragile]{}
    \frametitle{BPF program pack allocator}

    \begin{center}
        \begin{minted}[fontsize=\Large]{c}
    struct bpf_prog_pack {
        struct list_head list;
        void *ptr;
        unsigned long bitmap[];
    };

    // [...]

    ro_header = bpf_prog_pack_alloc(
                size, bpf_fill_ill_insns);
        \end{minted}
    \end{center}
\end{frame}
\note{
    57631054fae6dcc9c892ae6310b58bbb6f6e5048
    iTLB trashing due to small progs
}

\begin{frame}[fragile]{}
    \frametitle{Bloom filter map}

    \begin{center}
        \begin{minted}[fontsize=\Large]{c}
    bpf_map_create(
        BPF_MAP_TYPE_BLOOM_FILTER, NULL,
        0, sizeof(value), 100, NULL);
        \end{minted}
    \end{center}
\end{frame}
\note{
    9330986c03006ab1d33d243b7cfe598a7a3c1baa
}

\begin{frame}[fragile]{}
    \frametitle{Task local storage}

    \begin{center}
        \begin{minted}[fontsize=\Large]{c}
    ptr = bpf_task_storage_get(
            &start, t, 0,
            BPF_LOCAL_STORAGE_GET_F_CREATE);
        \end{minted}
    \end{center}
\end{frame}
\note{
    a10787e6d58c24b51e91c19c6d16c5da89fcaa4b
}

\fontsize{18pt}{18}\selectfont
\begin{frame}
  \vspace*{2.5cm}
  \begin{minipage}[b][\paperheight]{\textwidth}
  \begin{center}

      %\raggedright%
      \linespread{1.0}%
      \usebeamerfont{title}%
      \usebeamercolor[fg]{title}%
      \if@noSmallCapitals%
        Questions?
      \else%
        \scshape{\color{black} Questions?}%
      \fi%
      \vspace*{0.3em}

      \usebeamerfont{subtitle}%
      \fontsize{13pt}{14}\selectfont
      \usebeamercolor[fg]{subtitle}%
        \begin{itemize}[label={}]
            \item {\color{black} \twitter\ @erthalion}
            \item {\color{black} \email\ dmitrii.dolgov at redhat dot com}
        \end{itemize}
      \vspace*{2.5em}%

    \vfill
    \vspace*{2em}
  \end{center}
  \end{minipage}
\end{frame}

\end{document}